### 2022-03-13 proposal/proposal.tex
Initial Commit

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{4321}

\begin{document}

\title{Human Motion Interface for Video Game Control}
\maketitle

\section{Problem Description}
The human-computer interface behind many modern video games, using either handheld 
joystick controllers or keyboard and mouse input, is an example
of a typical human input device for computers, monitors, televisions, etc. There are two 
major drawbacks to traditional human input systems that are often overlooked: the input needed 
to achieve a certain effect has little to no correspondence to a physically-intuitive set 
of motions, and more importantly, inputs are heavily dependent on the user's fine motor skills. 
Fine motor skills used to interact with modern technology are heavily affected by factors such 
as muscular coordination disorders and old age. Computer vision, particularly the processes 
of motion tracking and human pose estimation, may allow us to create human-computer interfaces 
that link effect to physically-intuitive human motion inputs with little dependence on 
fine motor ability. We propose a vision-based computer-input system that maps gestures and 
motion of the user's body into computer input. We propose a car-racing simulator as a test 
subject for our system, and aim to intuitively control the motion of a player's virtual car 
with human body motion as the only action. Such a system would be instrumental in improving 
the accessibility of modern technology, with implementations far beyond the scope of computer 
games.

\section{Related Work}
\subsection{Human Pose Estimation}
Some of the largest components of our system fall into the category of so called Human 
Pose Estimation or the ability to properly:

\begin{itemize}
    \item Identify users as sources of input, from a video stream
    \item Segment the user's body to isolate relevant portions of data
    \item Compute location information about the user's body from relevant portions of data.
\end{itemize}

The area of Human Pose Estimation is widely studied. Robust, optimized solutions exist 
that can help us with this step. One popular solution is derived from \cite{8765346}, and 
open-source code for the so called openpose library shows promising results for real-time 
human body segmentation.

\begin{figure}[h]
    \centering
    \includegraphics[width = .8\linewidth]{openpose.png}
    \caption{A demonstration of openpose performing human body segmentation.}
\end{figure}

\subsection{Gesture Identification/Motion Tracking}
\cite{8765346} and Openpose give us promising ways to identify locations of human body features 
(hands, arms, etc.) in a video stream. However, this is not enough to categorize a user's motion 
as a particular input. To be able to extract feature motion over time as a useful input to our 
car racing video game, we will have to implement motion tracking over a video stream.

A number of guides exist for implementing feature tracking over video. \cite{tracking_1} 
and \cite{tracking_2} demonstrate simple object tracking across video frames, and \cite{tracking_2} 
demonstrates a system shown to be accurate with motion of hands. Implementing a similar system for 
tracking arbitrary features of the human body obtained from human pose estimation should provide 
adequate results.

\section{Methodology}
Our human computer interface will consist of three major steps: Pose Estimation, or feature 
location, Feature Tracking and classification of motions, and video game interfacing. Each one 
is detailed below. All code will be written in Python.

\subsection{Pose Estimation}
The first step of our human computer interface pipeline is the pose estimation step, in which 
we use openpose to locate certain key features of the user. What constitutes a key feature will 
be determined when the set of video game controls we must mimic is properly defined. Possible 
key features include the location of the user's wrists, elbows, or the angles of various limbs.

During this step we will also determine the computational cost of openpose and make decisions on 
how to proceed based on the results. If openpose is able to run with sufficiently low latency, we 
move on to feature tracking, but if openpose demonstrates a high latency, we will explore other 
methods of pose estimation, or methods of improving the performance of openpose using by decreasing 
framerates or image resolutions.

\subsection{Feature Tracking}
Obtaining the locations of various important features is only the first step towards identifying 
physically-meaningful actions of the user. In order for the user to be able to signal various 
commands within the video game, we must track the previously obtained features over time and 
classify the user's behavior into one of a few possible actions. For example, mimicing a car 
racing game, the set of possible actions could be \{No command, accelerate, decelerate, steer\}. 
Each of these commands or classifications will also have a value associated with it corresponding 
to its magnitude, which will also be obtained directly from feature locations. For example, a steering 
command mapped to the user rotating their torso would have a steering value associated with it 
derived from the angle through which the user has turned.

\subsection{Game Interface}
The third and final major step of our methodology is to turn the classifications obtained from 
the Feature Tracking step and feed them to a video game as user input. We will simply map each 
corresponding classification from step 2 to an "intermediate control" which is simply the 
joystick/keyboard/mouse input that would produce the same output from the game, and implement 
simple keyboard control with Python to feed that input to the video game.

\section{Timeline}
We adopt an eight-week timeline for the development and documentation of this project, outlined 
more specifically below:
\begin{itemize}
    \item Week 0: Configure openpose to identify feature points from front and/or side cameras.
    \item Week 1: Perform any necessary optimizations to run openpose with as low latency as possible.
    \item Week 2: Finalize the set of trackable and detectable gestures, and map each one to a particular keyboard mapping in a physically-intuitive way.
    \item Week 3-4: Implement a feature tracking system, and confirm that each gesture accurately corresponds to a specific function with sufficient accuracy.
    \item Week 5: Interface detections with video game.
    \item Week 6: Optimize vision video game interaction for low latency.
    \item Week 7: Write paper describing project methodology, and explore future works inspired by this project.
\end{itemize}

{\small
\bibliographystyle{ieee_fullname}
\bibliography{proposal_bib}
}

\end{document}

### 2022-03-14 proposal/proposal.tex
minor edits, LGTM

@@
 \section{Problem Description}
-The human-computer interface behind many modern video games, using either handheld 
-joystick controllers or keyboard and mouse input, is an example
-of a typical human input device for computers, monitors, televisions, etc. There are two 
+The human-computer interface behind many modern video games, uses either handheld 
+joystick controllers or keyboard and mouse input, which are examples
+of typical human input device for computers, monitors, televisions, etc. There are two 
 major drawbacks to traditional human input systems that are often overlooked: the input needed 
@@
 that link effect to physically-intuitive human motion inputs with little dependence on 
-fine motor ability. We propose a vision-based computer-input system that maps gestures and 
-motion of the user's body into computer input. We propose a car-racing simulator as a test 
+fine motor ability or additional physical input devices. We propose a vision-based computer-input system that maps gestures and 
+motion of the user's body into computer input. We also propose a car-racing simulator as a test 
 subject for our system, and aim to intuitively control the motion of a player's virtual car 
@@
 The area of Human Pose Estimation is widely studied. Robust, optimized solutions exist 
-that can help us with this step. One popular solution is derived from \cite{8765346}, and 
+that can help us with this step. One popular solution is derived from \cite{8765346} which uses Part Affinity Fields (PAFs) to learn to assosciate body parts with individuals in the image, and 
 open-source code for the so called openpose library shows promising results for real-time 
@@
     \item Week 2: Finalize the set of trackable and detectable gestures, and map each one to a particular keyboard mapping in a physically-intuitive way.
-    \item Week 3-4: Implement a feature tracking system, and confirm that each gesture accurately corresponds to a specific function with sufficient accuracy.
+    \item Week 3-4: Implement a real-time feature tracking system, and confirm that each gesture accurately corresponds to a specific function with sufficient accuracy.
     \item Week 5: Interface detections with video game.

### 2022-05-05 jetson_pose/detect_pose_dino.py
Improved comments and code quality in detect_pose_dino

@@ -1,31 +1,5 @@
-#!/usr/bin/python3
-#
-# Copyright (c) 2021, NVIDIA CORPORATION. All rights reserved.
-#
-# Permission is hereby granted, free of charge, to any person obtaining a
-# copy of this software and associated documentation files (the "Software"),
-# to deal in the Software without restriction, including without limitation
-# the rights to use, copy, modify, merge, publish, distribute, sublicense,
-# and/or sell copies of the Software, and to permit persons to whom the
-# Software is furnished to do so, subject to the following conditions:
-#
-# The above copyright notice and this permission notice shall be included in
-# all copies or substantial portions of the Software.
-#
-# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-# FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT.  IN NO EVENT SHALL
-# THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING
-# FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER
-# DEALINGS IN THE SOFTWARE.
-#
-
 import jetson.inference
 import jetson.utils
-
-import argparse
-import sys
 
 import numpy as np
 
@@ -35,11 +9,12 @@ keyboard = Controller()
 import matplotlib.pyplot as plt
 import time
 
-# load the pose estimation model
+# load the pose estimation model. The first time this runs it will 
+# Take some time to optimize the model for jetson TRT inference.
 net = jetson.inference.poseNet("resnet18-body", argv=["--log-level=error", f"--threshold={.15}"])
 
 # create video sources & outputs
-input = jetson.utils.videoSource("/dev/video0")
+cameraStream = jetson.utils.videoSource("/dev/video0")
 output = jetson.utils.videoOutput("display://0")
 
 # process frames until the user exits
@@ -52,23 +27,25 @@ prev_3_loc_index = 0
 THRESHOLD = 10 #px
 while (1):
     # capture the next image
-    img = input.Capture()
+    img = cameraStream.Capture()
 
[…]

### 2022-05-05 jetson_pose/detect_pose_dino.py
Improved comments and code quality for detect_pose_racer.py

@@ -3,6 +3,8 @@ import jetson.utils
 
 import numpy as np
 
+#pynput library for simulating keypresses
+#Controller() object for emulating keyboard
 from pynput.keyboard import Key, Controller
 keyboard = Controller()
 
@@ -14,6 +16,7 @@ import time
 net = jetson.inference.poseNet("resnet18-body", argv=["--log-level=error", f"--threshold={.15}"])
 
 # create video sources & outputs
+#/dev/video0 for USB Camera
 cameraStream = jetson.utils.videoSource("/dev/video0")
 output = jetson.utils.videoOutput("display://0")
 

### 2022-05-05 jetson_pose/detect_pose_racer.py

@@ -1,34 +1,10 @@
-#!/usr/bin/python3
-#
-# Copyright (c) 2021, NVIDIA CORPORATION. All rights reserved.
-#
-# Permission is hereby granted, free of charge, to any person obtaining a
-# copy of this software and associated documentation files (the "Software"),
-# to deal in the Software without restriction, including without limitation
-# the rights to use, copy, modify, merge, publish, distribute, sublicense,
-# and/or sell copies of the Software, and to permit persons to whom the
-# Software is furnished to do so, subject to the following conditions:
-#
-# The above copyright notice and this permission notice shall be included in
-# all copies or substantial portions of the Software.
-#
-# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-# FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT.  IN NO EVENT SHALL
-# THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING
-# FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER
-# DEALINGS IN THE SOFTWARE.
-#
-
 import jetson.inference
 import jetson.utils
 
-import argparse
-import sys
-
 import numpy as np
 
+#pynput library for simulating keypresses
+#Controller() object for emulating keyboard
 from pynput.keyboard import Key, Controller
 keyboard = Controller()
 
@@ -36,53 +12,62 @@ keyboard = Controller()
 net = jetson.inference.poseNet("resnet18-body", argv=["--log-level=error", f"--threshold={.15}"])
 
 # create video sources & outputs
-input = jetson.utils.videoSource("/dev/video0")
+# /dev/video0 for USB Camera
+videoStream = jetson.utils.videoSource("/dev/video0")
 output = jetson.utils.videoOutput("display://0")
+
+#Experimentally tuned cuttoff
+TURNING_CUTOFF = .20
 
 # process frames until the user exits
 while True:
     # capture the next image
-    img = input.Capture()
+    img = videoStream.Capture()
 
-    # perform pose estimation (with overlay)
+    # perform pose estimation (with keypoint overlay)
[…]

### 2022-05-07 paper/paper.tex
included final paper template
Abstract and introduction

--- 2022-03-14 proposal/proposal.tex
@@
 %\ifcvprfinal\pagestyle{empty}\fi
-\setcounter{page}{4321}
+\setcounter{page}{1}
 
@@
 
-\title{Human Motion Interface for Video Game Control}
+\title{Human Pose Estimation for Video Game Control}
+\author{Rakesh Johny, Tom Li, Aditya Narayanan, Albert Xia}
 \maketitle
 
-\section{Problem Description}
-The human-computer interface behind many modern video games, uses either handheld 
+\begin{abstract}
+    Current methods of interacting with computers are flawed in a 
+    couple of key ways: they fail to map physically-intuitive motions to their computer 
+    control counterparts, and they rely heavily on a user's fine-motor skills, 
+    which are heavily impacted by factors such as muscle coordination disabilities 
+    and old age. In this paper, we propose a system of computer control through 
+    gesture tracking via real-time human pose estimation on an embedded device, 
+    which is capable of addresssing these issues, by mapping general physically-intuitive 
+    motions into computer control. Furthermore, our device boasts low latency in 
+    the human-computer interaction, and is minimally intrustive to the computer being 
+    controlled. We test the viability of our system as a computer-control device by playing two 
+    video games using only gestures.
+\end{abstract}
+
+\section{Introduction}
+The human-computer interfaces behind many modern video games use either handheld 
 joystick controllers or keyboard and mouse input, which are examples
@@
 as muscular coordination disorders and old age. Computer vision, particularly the processes 
-of motion tracking and human pose estimation, may allow us to create human-computer interfaces 
-that link effect to physically-intuitive human motion inputs with little dependence on 
-fine motor ability or additional physical input devices. We propose a vision-based computer-input system that maps gestures and 
-motion of the user's body into computer input. We also propose a car-racing simulator as a test 
-subject for our system, and aim to intuitively control the motion of a player's virtual car 
-with human body motion as the only action. Such a system would be instrumental in improving 
-the accessibility of modern technology, with implementations far beyond the scope of computer 
-games.
+of human pose estimation and motion tracking may allow us to create human-computer interfaces 
+that link computer control to physically-intuitive human motion inputs with little dependence on 
+fine motor ability or additional physical input devices. We propose a vision-based computer-input 
+system that maps gestures and motion of the user's body into computer input. We test the system's 
+effectiveness with the playability of two simple video games as metrics. In order to demonstrate 
+the versatility of our phased-approach to gesture-based control, we select a car-racing game and 
+the Google Chrome Dino Racer game as test video games. Such a system, when expanded to replicating 
+general keyboard + mouse control, would be instrumental in improving the accessibility of 
+modern technology, with implementations far beyond the scope of computer games.
 
@@
     \centering
-    \includegraphics[width = .8\linewidth]{openpose.png}
+    \includegraphics[width = .8\linewidth]{images/openpose.png}
     \caption{A demonstration of openpose performing human body segmentation.}
@@
 
-\section{Timeline}
-We adopt an eight-week timeline for the development and documentation of this project, outlined 
-more specifically below:
-\begin{itemize}
-    \item Week 0: Configure openpose to identify feature points from front and/or side cameras.
-    \item Week 1: Perform any necessary optimizations to run openpose with as low latency as possible.
-    \item Week 2: Finalize the set of trackable and detectable gestures, and map each one to a particular keyboard mapping in a physically-intuitive way.
-    \item Week 3-4: Implement a real-time feature tracking system, and confirm that each gesture accurately corresponds to a specific function with sufficient accuracy.
-    \item Week 5: Interface detections with video game.
-    \item Week 6: Optimize vision video game interaction for low latency.
-    \item Week 7: Write paper describing project methodology, and explore future works inspired by this project.
-\end{itemize}
+\section{Results}
+
+\section{Future Works}
 

### 2022-05-07 paper/paper.tex
Finished up the related works section

@@
     \item Segment the user's body to isolate relevant portions of data
-    \item Compute location information about the user's body from relevant portions of data.
+    \item Compute the location, in image coordinates, of keypoints on the user's body.
 \end{itemize}
 
-The area of Human Pose Estimation is widely studied. Robust, optimized solutions exist 
-that can help us with this step. One popular solution is derived from \cite{8765346} which uses Part Affinity Fields (PAFs) to learn to assosciate body parts with individuals in the image, and 
-open-source code for the so called openpose library shows promising results for real-time 
+The area of Human Pose Estimation is widely studied. There are a wide array of robust, optimized 
+solutions for gathering 2D image coordinates of keypoints from images of users.
+
+\subsubsection{Openpose}
+One popular solution to realtime human pose estimation, derived from \cite{8765346}, 
+uses Part Affinity Fields (PAFs) to learn to assosciate body parts with individuals in the image, 
+and open-source code for the so called openpose library shows promising results for real-time 
 human body segmentation.
@@
 
+\subsubsection{MoveNet and PoseNet}
+
+PoseNet and its more recent counterpart MoveNet are fully convolutional neural network pose 
+estimators built on existing ResNet or MobileNet backbones (citation here). Benchmarks of PoseNet and MoveNet show that it is 
+viable to run either estimator in real time on embedded devices with appropriate 
+model quantizations (citation here), with PoseNet being slightly faster when compared to MoveNet at identical 
+levels of quantization. In comparison's with Openpose, PoseNet/MoveNet are shown to perform 
+slightly better on instance segmentation of the person class from the COCO dataset.
+
 \subsection{Gesture Identification/Motion Tracking}
-\cite{8765346} and Openpose give us promising ways to identify locations of human body features 
+Openpose and MoveNet/PoseNet give us promising ways to identify locations of human body features 
 (hands, arms, etc.) in a video stream. However, this is not enough to categorize a user's motion 
-as a particular input. To be able to extract feature motion over time as a useful input to our 
-car racing video game, we will have to implement motion tracking over a video stream.
+as a particular command to a computer system. To be able to extract feature motion over time 
+as a useful input to our car racing video game, we will have to implement motion tracking 
+over a video stream.\\
 
@@
 and \cite{tracking_2} demonstrate simple object tracking across video frames, and \cite{tracking_2} 
-demonstrates a system shown to be accurate with motion of hands. Implementing a similar system for 
-tracking arbitrary features of the human body obtained from human pose estimation should provide 
-adequate results.
+demonstrates a system shown to be accurate with motion of hands. We drew on \cite{tracking_1} and 
+\cite{tracking_2} in developing a tracking algorithm for the identification of a human user ducking 
+or jumping from a keypoint cloud.
 

### 2022-05-07 paper/paper.tex
Finished pose estimation section

@@
 \section{Methodology}
-Our human computer interface will consist of three major steps: Pose Estimation, or feature 
-location, Feature Tracking and classification of motions, and video game interfacing. Each one 
-is detailed below. All code will be written in Python.
+Our human computer interface consists of three major steps: Pose Estimation, or feature 
+location, Feature Tracking and classification of motions, and video game interfacing. In addition, 
+we adopt a multi-phased approach to our human computer interface. Phase 1 consists of a robust and 
+general human pose estimator and keypoint detector. The work in Phase 1 is task/game agnostic, as 
+we implement identical versions of this phase for the car-racer game as well as the chrome 
+dinosaur game. Phase 2 is a game-specific keypoint classifier, where the output from Phase 1 is 
+classified into a number of computer commands in a way dependent on the physical motions being 
+replicated in the game as well as the computer commands the game is traditionally played with.\\
+
+The motivation behind such a two phase approach rather than building a classifier that directly 
+maps input video streams to computer commands is to improve the versatility of our human-computer 
+interaction system. A fully general Phase 1 implemented independent of the game in question allows 
+the system to be more easily adapted to any particular set of computer commands. We demonstrate the 
+versatility of this system through the implementation of our human-computer interface for two 
+distinct video games that utilize identical Phase 1 code.
+
+\subsection{Testing Pose Estimation Methods}
+Before we develop phase 1 of our human-computer interface, we test a number of proposed solutions 
+for human pose estimation to determine which ones are best-suited to our needs. The main criteria 
+we require for our human pose estimator are:
+
+\begin{itemize}
+    \item Real time operation. The need for this criterion is obvious, as we require low latency 
+        classification in order to mimic low-latency commmands for video games. Essentially, the user 
+        should feel as if the delay between executing a physical command and the video game reflecting 
+        such a command is instantaneous.
+    \item The system should be minimally intrusive on the computer used to play the game. This 
+        essentially means that we cannot use computing resources so significant that usability of 
+        the computer itself is diminished. 
+\end{itemize}
+
+\subsubsection{Openpose Python API}
+The openpose package offers a Python API for implementing inference over video feeds using one 
+of the openpose models. We began testing the openpose Python API for our task, but found that 
+installing the API resulted in dependency issues accross multiple team members' computers and 
+decided to instead test the openpose pretrained models implemented independently of the openpose 
+python API.
+
+%https://learnopencv.com/deep-learning-based-human-pose-estimation-using-opencv-cpp-python/
+\subsubsection{Openpose model implemented in OpenCV}
+(citation) offers a method to perform inference on CPU/GPU using pretrained versions of the 
+openpose hand/body pose estimators. For inference with the openpose pretrained model through 
+OpenCV, we observed only 1-2 FPS on a 2.3 GHz Intel i9 CPU. We expect that running the same model 
+with CUDA acceleration would result in true real-time performance, but as one of our design 
+constraints for our system, we imposed that we cannot use significant GPU resources in order to 
+maintain minimal intrusiveness onto the gaming machine.
+
+\subsubsection{PoseNet and the Jetson Nano}
+Due to the poor performance of the Openpose models, we began to look at different methods for 
+real-time pose estimation. One option that was promising was to completely offload the task of 
+human pose estimation to an edge-device to reduce the computational burden on the device used 
+to play the games. Such an embedded device would ideally be compact yet powerful enough to run 
+quantized versions of pose estimators in real time. We found the NVIDIA Jetson Nano to be an 
+ideal candidate for such an embedded device.\\
+
+With a 128-core NVIDIA GPU on a single-board computer footprint, the Jetson Nano is a popular 
+choice for deep model inference on embedded computers. Perhaps more important than the hardware 
+itself, however, is the specific Tensor RT optimized models available for use on the Jetson Nano. 
+Through a number of optimization techniques such as model quantization, unuzed output elimination 
+and horizontal layer fusion/layer aggregation, the Tensorflow with Tensor RT (TF-TRT) backend is 
+able to greatly improve inference time on the Jetson Nano. TF-TRT is able to output optimized 
+models from Open Neural Network Exchange Format (ONXX) and for popular tasks such as human pose 
+estimation, ONXX model formats are readily available from NVIDIA Developers.
 
 \subsection{Pose Estimation}
-The first step of our human computer interface pipeline is the pose estimation step, in which 
-we use openpose to locate certain key features of the user. What constitutes a key feature will 
-be determined when the set of video game controls we must mimic is properly defined. Possible 
-key features include the location of the user's wrists, elbows, or the angles of various limbs.
+Using a variant of PoseNet built on a ResNet-18 backbone, we observe 15-17 Frames Per Second 
+(FPS) on the Jetson Nano. The so called Pose-ResNet-18-body outputs 17 detected keypoints in 
+image coordinate format. 15-17 FPS performance is smooth enough keypoint detection to allow us to 
+play these simple games in real time.
 
-During this step we will also determine the computational cost of openpose and make decisions on 
-how to proceed based on the results. If openpose is able to run with sufficiently low latency, we 
-move on to feature tracking, but if openpose demonstrates a high latency, we will explore other 
-methods of pose estimation, or methods of improving the performance of openpose using by decreasing 
-framerates or image resolutions.
-
-\subsection{Feature Tracking}
-Obtaining the locations of various important features is only the first step towards identifying 
-physically-meaningful actions of the user. In order for the user to be able to signal various 
-commands within the video game, we must track the previously obtained features over time and 
-classify the user's behavior into one of a few possible actions. For example, mimicing a car 
-racing game, the set of possible actions could be \{No command, accelerate, decelerate, steer\}. 
-Each of these commands or classifications will also have a value associated with it corresponding 
-to its magnitude, which will also be obtained directly from feature locations. For example, a steering 
-command mapped to the user rotating their torso would have a steering value associated with it 
-derived from the angle through which the user has turned.
+\subsection{Feature Tracking and Classification}
+\subsubsection{Feature Classification For Steering}
+\subsubsection{Feature Classification For Jumping/Ducking}
 
 \subsection{Game Interface}
-The third and final major step of our methodology is to turn the classifications obtained from 
-the Feature Tracking step and feed them to a video game as user input. We will simply map each 
-corresponding classification from step 2 to an "intermediate control" which is simply the 
-joystick/keyboard/mouse input that would produce the same output from the game, and implement 
-simple keyboard control with Python to feed that input to the video game.
 
@@
 
-\section{Future Works}
+\section{Conclusion and Future Works}
 

### 2022-05-07 paper/paper.tex
Added feature classification for steering

@@
 
-With a 128-core NVIDIA GPU on a single-board computer footprint, the Jetson Nano is a popular 
+With a 128-core NVIDIA GPU on a single-board footprint, the Jetson Nano is a popular 
 choice for deep model inference on embedded computers. Perhaps more important than the hardware 
@@
 models from Open Neural Network Exchange Format (ONXX) and for popular tasks such as human pose 
-estimation, ONXX model formats are readily available from NVIDIA Developers.
+estimation, ONXX model formats are readily available from NVIDIA Developers.\\
 
-\subsection{Pose Estimation}
 Using a variant of PoseNet built on a ResNet-18 backbone, we observe 15-17 Frames Per Second 
@@
 image coordinate format. 15-17 FPS performance is smooth enough keypoint detection to allow us to 
-play these simple games in real time.
+play games in real time.
+
+\subsection{Pose Estimation}
+After testing the variety of available solutions for human pose estimation 
+above, we identify that running TRT optimized PoseNet through the Jetson Nano 
+is the best solution for our application.\\
+
+To build the first phase of our human-computer interaction device, we use the 
+jetson-inference and jetson-utils python packages (citation here) 
+to build a human pose estimator in Python. The output of this step is a 
+collection of 2D keypoints, correpsonding to image coordinates of the locations 
+of the 17 body parts detected by PoseNet.
 
 \subsection{Feature Tracking and Classification}
+The second major phase of the human-computer interaction device are the 
+keypoint classifiers, which receive 2d image coordinates keypoint detections 
+from the pose estimator, and classify collections of keypoints into computer 
+commands in a game-specific way. In a fully generalized model of our 
+human-computer interaction device, the keypoint "decoder" built in this step 
+need not even be a classifier. For more complicated continuous inputs to 
+video games, this can be a more complicated regression problem as we would 
+seek to estimate continuous output based on the locations of human pose 
+keypoints.\\
+
+We construct two distinct keypoint classifiers, one for each video game we 
+demonstrate our system on. We construct one classifier to derive a steering 
+command from pose keypoints, and one to derive jumping/ducking information 
+from an identical set of keypoints.
+
 \subsubsection{Feature Classification For Steering}
+The first of two example video games we play as demonstrations of our 
+human-computer interface is a simple car-racing game, in which the player's 
+vehicle must dodge obstacles by steering left and right. The video game 
+accepts controls as either left arrow key or right arrow key, so our 
+keypoint classifier for this game seeks to classify collections of keypoints 
+derived from our human pose estimator into either steer left or steer right 
+commands.\\
+
+To maintain a physical intuition to the gestures that are mapped to either 
+left arrow key or right arrow key, we require a user to steer the car with 
+an imaginary steering wheel on the video stream. From each frame of the user's 
+hands on an imaginary wheel, we extract classifications for the game control 
+as follows:
+
+\begin{enumerate}
+    \item Assert that both the left wrist and right wrist of the user on the 
+        imaginary wheel are present in the frame. If one or more wrists are 
+        not present in the frame, skip the remaining steps and output no 
+        steering command.
+    \item Compute the vector connecting the left wrist of the user to the 
+        right wrist of the user.
+    \item Compute the angle to the horizontal of the vector computed in step 2.
+    \item Threshold on the angle to the horizontal with an experimentally 
+        determined cutoff threshold.
+\end{enumerate}
+
+Let the location (in image coordinates) of the user's left wrist be $(x,y)$ and 
+the location of the user's right wrist be $(x', y')$. The displacement vector 
+connecting the two wrists is given by:
+
+$$d = (x - x', y - y')$$
+
+The angle to the horizontal is then given by:
+
+$$\theta = \arctan{\frac{y-y'}{x-x'}}$$
+
+After some experimentation with the system, we found that a reasonable threshold 
+for the steering angle is $0.2$ radians. Our classifications for steering angle 
+are then derived as:
+
+\begin{itemize}
+    \item $\theta > 0.2$: Steer left
+    \item $-0.2 < \theta < 0.2$: No command
+    \item $-0.2 < \theta$: Steer Right
+\end{itemize}
+
 \subsubsection{Feature Classification For Jumping/Ducking}

### 2022-05-08 paper/paper.tex
Added feature tracking for dino game

@@
 \subsubsection{Feature Classification For Jumping/Ducking} 
+The second example video game we demonstrate our human-computer interface system 
+on is the Google Chrome Dinosaur Game. The game accepts jumping and ducking commands 
+in the form of up arrow key and down arrow key; however, our classifier is a 
+little more complex than for the first example video game, as ducking and jumping 
+require motion tracking over video to properly classify from human motion. That is, 
+you cannot determine from a single frame of video whether or not the user is 
+ducking or jumping.\\
+
+In order to deisgn an appropriate classifier for this task, we first establish 
+a subset of the keypoints as a more descriptive feature to track. In the case of 
+the car steering example, this was done by reducing the set of keypoints to simply 
+the left wrist and right wrist. For classification of jumping and ducking, this is 
+accomplished by computing a "mean face" or the midpoint of a number of selected 
+keypoints. For this application, we select the two eyes, two ears, and the nose, 
+and compute the midpoint of these keypoints as our augmented feature to track in 
+order to classify whether a user is jumping or ducking. The motivation for establishing 
+this feature is to improve the robustness of our classification. Were we to simply 
+track a single keypoint such as the nose of the user, our system would not be 
+stable against random errors when the nose may not be properly classified or 
+may not be classified at all from the phase 1 human pose estimation. Running 
+relatively small versions of PoseNet, these types of errors are fairly common. 
+Establishing the midpoint of the face as a feature to track allows us to track 
+an equally useful but more robust feature.\\
+
+After establishing the midpoint feature, we obtain graphs of the vertical 
+position and derivative of vertical position of the midpoint feature over time 
+as a user performs jumping and ducking motions in order to develop a game-specific 
+classifier for this task. A few methods we considered and our analysis of them 
+are given below:
+
+\begin{enumerate}
+    \item \textbf{Thresholding on Position}: The most simple method we could 
+        employ to classify the movement of the midpoint face into either jumping 
+        or ducking would be to threshold on the vertical position in the frame 
+        of the midpoint of the users's face. However, this system is not invariant 
+        to small motions of the user between jumps and ducks. For example, if the 
+        user moves closer to or farther from the camera, this method breaks down.
+    \item \textbf{Threshold on Derivative}: One method we considered would be 
+        to simply threshold on the derivative of the y-position to discriminate 
+        between periods of slow/no motion, and those where the user rapidly moves 
+        up or down as one would during a jump or a duck. However, as we can see from 
+        the graphs of the derivative over time, both jumps and ducks have periods of 
+        high derivative in both the positive and negative direction. i.e) thresholding 
+        on the deriviatve is not enough to discriminate between a duck and a jump, although 
+        it can discriminate between jump/duck and no command.
+    \item \textbf{Template Matching}: From the graph of the derivative over time, we can 
+        see that a jump and a duck have inverted profiles on the graph. If we compute 
+        the correlation between a template graph for a jump/duck and the current 
+        graph at all timesteps, we would be able to discriminate between a jump and 
+        a duck effectively. However, this method cannot be used in real time. As we can 
+        see from either graph, a jump/duck occupies a significant portion of time. Computing 
+        the correlation between a template feature and the motion of the midpoint face 
+        in real time would only result in classification at the end of a jump or duck, which 
+        would introduct unwanted latency into the system.
+    \item \textbf{Threshold on Derivative with extra constraint}: The 
+        method we implement in our system is a modification of simply thresholding on 
+        the derivative of the y position. We threshold on the vertical position, with 
+        the added constraint that at any time $t$, for the sample at that time to be 
+        classified as a jump or a duck, the previous $N$ samples of the 
+        derivative must be below the threshold. This ensures that only the 
+        first period of rapid change in the motion of the midpoint face is 
+        captured as a command. $N$ and the threshold are tunable parameters, 
+        and in our final system, we set the threshold = $10$ pixels/sec and 
+        $N = 3$.
+\end{enumerate}
+
+The final solution we implement for our feature tracking/classification for 
+the Chrome dino game is as follows:
+
+\begin{enumerate}
+    \item For every frame of video, compute $y_t$, the midpoint of all keypoints 
+        existing in the frame from the set 
+        \{left eye, right eye, left ear, right ear, nose\}.
+    \item Compute $y'_t$, the difference of $y_t$ and $y_{t-1}$.
+    \item If $y'_t > 10, y'_{t-1} < 10, y'_{t-2} < 10$ and $y'_{t-3} < 10$: 
+        Classify current frame as a duck.
+    \item If $y'_t < 10, y'_{t-1} < 10, y'_{t-2} < 10$ and $y'_{t-3} < 10$:
+        Classify current frame as a jump.
+    \item If none of the above conditions are satisfied, classify current frame as 
+        no command.
+\end{enumerate}
+
+\textit{Note that $y' > 10$ corresponds to a duck and $y' < 10$ corresponds 
+to a jump as the image coordinates are zeroed at the top left of the frame.}
 

### 2022-05-08 paper/paper.tex
minor typo change in feature tracking for chrome dino game
added game interface section

@@
 \begin{enumerate}
-    \item For every frame of video, compute $y_t$, the midpoint of all keypoints 
+    \item For every frame of video, compute $y_t$, the midpoint of y-components 
+    of all keypoints 
         existing in the frame from the set 
@@
 \subsection{Game Interface}
+The final step before our human-computer interaction system is fully functional 
+is to simulate keypresses to play the two example games selected. In this project 
+we select simple example games that can be run on the Jetson Nano within a browser 
+or through pyGame. However, the system was designed so that during this phase, the 
+Jetson Nano will send data to a game console machine either through serial 
+communication or wirelessly through bluetooth/wifi communication to play more 
+intensive games on a dedicated machine.\\
+
+In order to simulate keypresses, we utilize a Python package known as pynput, 
+which allows us to press and release arrow keys at the appropriate times in order 
+to simulate keypresses to the example video games.
 

### 2022-05-08 paper/paper.tex
Finished up conclusion and future works

@@
 or through pyGame. However, the system was designed so that during this phase, the 
-Jetson Nano will send data to a game console machine either through serial 
+Jetson Nano could send data to a game console machine either through serial 
 communication or wirelessly through bluetooth/wifi communication to play more 
@@
 \section{Results}
+Once the human pose estimator, keypoint classifiers, and game interface are integrated into 
+a single system, we are able to play both of our example games in real time (15+ FPS) through only our 
+gestures in front of the camera. Our systems are robust, performing well across a variety of scenarios, 
+including when keypoints are missing and erroneous pose estimation. The games are responsive and 
+fully playable from only gestural input, demonstrating little to no loss of playability when 
+moving from keyboard and mouse input to the gestural command input through our human-computer 
+interface device.\\
 
 \section{Conclusion and Future Works}
+The goal of this project was to develop a system that can map human gestures through a video 
+stream into usable computer command input, replacing how we currently interact with computers 
+through keyboards, mice, and joysticks. Furthermore, we aimed to develop a system that 
+places little to no computational expense on the computer playing the game. 
+Our system demonstrates that we accomplish this goal 
+in real time, by playing two example video games using gestural inputs, with computation 
+and realtime inference done on an edge device, the NVIDIA Jetson Nano.\\
+
+In developing and testing this system, we note a few major areas in which the system can 
+be improved in the future. We notice that pose estimations using PoseNet on the Jetson nano 
+fails to detect keypoints around the wrist dependent on the position and condition of the user's 
+hands. We notice that detection is more stable over time when the user's hands are open rather than in the 
+closed fist position, although keypoints are still correctly identified in both positions. One 
+area of future work that can address this problem is to retrain pose estimation models using a 
+custom dataset with more examples of situations such as mimicing steering motions.\\
+
+Other areas of future work that can improve our system and bring it closer to the goal of a 
+general purpose human computer interface that can mimic a wider range of commands include more 
+general keypoint classifiers using machine learning techniques that can be easily expanded 
+to a wider range of commands. In addition, a communication link between the Jetson Nano and 
+a gaming device can be implemented, allowing the user to play games on other devices, using our 
+human-computer interface only as an intermediate step.
 

### 2022-05-08 paper/paper.tex
added missing citations

@@
 PoseNet and its more recent counterpart MoveNet are fully convolutional neural network pose 
-estimators built on existing ResNet or MobileNet backbones (citation here). Benchmarks of PoseNet and MoveNet show that it is 
+estimators built on existing ResNet or MobileNet backbones \cite{PoseNet}. Benchmarks of PoseNet and MoveNet show that it is 
 viable to run either estimator in real time on embedded devices with appropriate 
-model quantizations (citation here), with PoseNet being slightly faster when compared to MoveNet at identical 
+model quantizations \cite{tensorflowPoseNet}, with PoseNet being slightly faster when compared to MoveNet at identical 
 levels of quantization. In comparison's with Openpose, PoseNet/MoveNet are shown to perform 
@@
 \subsubsection{Openpose model implemented in OpenCV}
-(citation) offers a method to perform inference on CPU/GPU using pretrained versions of the 
+\cite{opencvPose} offers a method to perform inference on CPU/GPU using pretrained versions of the 
 openpose hand/body pose estimators. For inference with the openpose pretrained model through 
@@
 To build the first phase of our human-computer interaction device, we use the 
-jetson-inference and jetson-utils python packages (citation here) 
+jetson-inference and jetson-utils python packages (\cite{jetsonInference}) 
 to build a human pose estimator in Python. The output of this step is a 
@@
 \bibliographystyle{ieee_fullname}
-\bibliography{proposal_bib}
+\bibliography{report_bib}
 }

### 2022-05-08 paper/paper.tex
Added a couple of images

@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width = .8\linewidth]{images/flow_chart.png}
+    \caption{Data flow within our human-computer interface system. Note 
+    the use of a universal pose estimator and game-specific classifiers.}
+\end{figure}
+
 \subsection{Testing Pose Estimation Methods}
@@
 collection of 2D keypoints, correpsonding to image coordinates of the locations 
-of the 17 body parts detected by PoseNet.
+of the 17 body parts detected by PoseNet.\\
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width = .8\linewidth]{images/pose_estimations.jpg}
+    \caption{An example of the keypoints detected from a frame of video
+    by Pose-ResNet-18}
+\end{figure}
 

### 2022-05-09 paper/paper.tex
added graphs for motion tracking
edits, added a gameplay image

@@
     Current methods of interacting with computers are flawed in a 
-    couple of key ways: they fail to map physically-intuitive motions to their computer 
+    couple of key ways, especially when it comes to video games: they fail to map physically-intuitive motions to their computer 
     control counterparts, and they rely heavily on a user's fine-motor skills, 
@@
 and open-source code for the so called openpose library shows promising results for real-time 
-human body segmentation.
+multi-person 2D pose detection, including body, foot, hand, and facial keypoints.
 
@@
 and \cite{tracking_2} demonstrate simple object tracking across video frames, and \cite{tracking_2} 
-demonstrates a system shown to be accurate with motion of hands. We drew on \cite{tracking_1} and 
+demonstrates a system shown to be accurate with motion of hands.This is done by calculating the changes between consecutive frames and finding trends in a window of input data.
+We drew on \cite{tracking_1} and
 \cite{tracking_2} in developing a tracking algorithm for the identification of a human user ducking 
@@
 classifier for this task. A few methods we considered and our analysis of them 
-are given below:
+are given below.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width = .8\linewidth]{images/center_height.png}
+    \includegraphics[width = .8\linewidth]{images/center_deriv.png}
+    \caption{Top: y-position of face midpoint over time. Bottom: derivative 
+    of y-position of face midpoint over time.}
+\end{figure}
 
@@
 interface device.\\
-
+\begin{figure}[h]
+    \centering
+    \includegraphics[width = .5\linewidth,height=4cm]{images/user.png}\hfill
+    \includegraphics[width = .5\linewidth,height=4cm]{images/game.png}
+    \caption{Real-time game play using our interface}
+\end{figure}
 \section{Conclusion and Future Works}
@@
 general keypoint classifiers using machine learning techniques that can be easily expanded 
-to a wider range of commands. In addition, a communication link between the Jetson Nano and 
+to a wider range of commands. As an immediate next step, we could develop kepoint classifiers that are specific to a video game genre (such as RPGs or shooters) rather than a sepcific game.
+In addition, a communication link between the Jetson Nano and 
 a gaming device can be implemented, allowing the user to play games on other devices, using our 

### 2022-05-09 paper/paper.tex
Added another figure and fixed some typos

@@
     couple of key ways, especially when it comes to video games: they fail to map physically-intuitive motions to their computer 
-    control counterparts, and they rely heavily on a user's fine-motor skills, 
+    control counterparts, and they rely on a user's fine-motor skills, 
     which are heavily impacted by factors such as muscle coordination disabilities 
@@
 as muscular coordination disorders and old age. Computer vision, particularly the processes 
-of human pose estimation and motion tracking may allow us to create human-computer interfaces 
+of human pose estimation and motion tracking, may allow us to create human-computer interfaces 
 that link computer control to physically-intuitive human motion inputs with little dependence on 
@@
 estimation, ONXX model formats are readily available from NVIDIA Developers.\\
+
+\begin{figure}
+    \centering
+    \includegraphics[width = .6\linewidth]{images/jetson_nano.jpg}
+    \caption{NVIDIA Jetson Nano, a single-board computer capable of 
+    low-latency deep model inference.}
+\end{figure}
 
